---
title: "PML-Training"
author: "Torsten Glunde"
date: "13 Aug 2014"
output: html_document
---

Background <http://groupware.les.inf.puc-rio.br/har>.

Data being used: <https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv>

First we are loading the data and do some exploration.

```{r}
if(!file.exists("training.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", method = "curl", destfile="training.csv")  
}
if(!exists('training')) training <- read.csv("training.csv")
dim(training)
colnames(training)
training$classe <- as.factor(training$classe)
classe <- as.factor(training$classe)
```

* We ignore the fact exercises are being executed by different persons and also the time variance - that people might get tired doing the exercises one after each other.
* We remove all features that are non-numeric, have NAs or 

```{r}
tmp <- training
tmp <- tmp[,which(as.numeric(colSums(is.na(tmp)))==0)]
tmp <- tmp[,which(sapply(tmp,class)!='factor')]
tmp <- subset(tmp, select=-X)
tmp <- subset(tmp, select=-raw_timestamp_part_1)
tmp <- subset(tmp, select=-raw_timestamp_part_2)
tmp <- subset(tmp, select=-num_window)
train <- cbind(classe,tmp)

colnames(train)
summary(train)
```

Finally we are fitting the random forest model looking at feature importance:
```{r}
require(randomForest)
set.seed(4242)
rfMDL <- randomForest(classe~., data=train, na.action=na.fail, ntree=502)
print(rfMDL)
par(mfrow=c(1,1)) ; varImpPlot(rfMDL,sort=TRUE)
```

And we can predict using the test data:
```{r}
set.seed(4242)
if(!file.exists("pml-testing.csv")) {
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", method = "curl", destfile="pml-testing.csv")  
}
if(!exists('testing')) testing <- read.csv("pml-testing.csv")

prediction <- predict(rfMDL, testing)
prediction
```
